Preliminary Draft

### 2021-07-20 article/draft.tex
small date precision for YT flags and Youtube -> YouTube

Section $230$ in the United States Communications Decency Act provides immunity for website platforms against the content created by users. Similar regulations exist in the European Union via the E-commerce Directive (2000) in articles $12$ and $15$.\footnote{See section $4$ in Bayer (2019)~\cite{Bayer} for a comprehensive overview of the mentioned articles.} Nevertheless, there is growing pressure for Mainstream Social Media Platforms, such as Facebook, Twitter or Youtube, to moderate the available content. In particular, platforms take explicit actions when content is in violation of local laws in different jurisdictions, such as laws regarding defamation of a racial nature, dissemination of symbols from unconstitutional organizations, privacy protection, digital security, electoral laws. For example, Facebook reports having implemented a total of $64.7$ thousand content restrictions based on local law across all countries in 2020.\footnote{See Facebook Transparency Center, Content restrictions based on Local Law: \href{https://transparency.fb.com/data/content\-restrictions}{transparency.fb.com/data/content\-restrictions}. We summed the count of content restrictions over all countries reported in the table, for $H1$ and $H2$ of the year 2020.} %Google reports a total of $26$ thousand government requests to remove content from July 2020 to December 2020, among which $11.4$ thousand concerned Youtube.\footnote{See Google's Transparency report, government requests to remove content: \href{https://transparencyreport.google.com/government-removals/overview}{transparencyreport.google.com/government-removals/overview}.} Twitter reports having received $42.2$ thousand legal demands from third-parties from January to June 2020, and has responded by withholding $82$ thousand accounts and $3.1$ thousand tweets.\footnote{See Twitter Transparency website, Removal requests: \href{https://transparency.twitter.com/en/reports/removal-requests.html\#2020-jan-jun}{transparency.twitter.com/en/reports/removal-requests.html\#2020-jan-jun}.}

@@

Furthermore, mainstream platforms are increasingly engaging in editorial tasks by implementing targeted policies to insure that each platform's rules are not violated. Community guidelines of Facebook, Twitter and Youtube can be summarized in a handful of categories, regarding safety, privacy and authenticity; which include sub-categories such as violence, terrorism, child sexual exploitation, abuse, harassment, hateful conduct, suicide or self-harm, illegal or regulated goods and services, platform manipulation and spam (see table \ref{summary} in Appendix \ref{links}). While specific to each platform, the previously cited categories correspond in most cases to well defined concepts that fall into legal frameworks in many countries, unlike misinformation. The intricacies of constructing a legal framework for misinformation arises from the difficulty of identifying and qualifying a piece of online content as false or misleading, among an overwhelming quantity of daily produced content, without infringing existing laws.\footnote{See \href{https://www.poynter.org/ifcn/anti-misinformation-actions/}{A guide to anti-misinformation actions around the world} on the website of Poynter Institute.} In particular, a number of recent studies  point towards the idea that ``Fake News'' or disinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense. \footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.'' Furthermore, the European commission recognizes the spread of online disinformation as a problem and has put together in 2018 a \href{https://digital-strategy.ec.europa.eu/en/policies/code-practice-disinformation}{Code of practice on Disinformation}, which is a set of self-regulatory standards to fight disinformation.}

@@

Hence, in the present article, we focus on mainstream platforms' policies and interventions regarding content with low credibility or false information, commonly referred to as {\it Fake News} (see Lazer et al. (2018)~\cite{lazer}). The {\it Fake News} phenomenon is still ill-defined by the academic community, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience.  The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society, in the rare event that it goes viral. This gave rise to a set of heterogenous fact-checking policies across mainstream platforms. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network. Facebook uses a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''}} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation (see \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{Twitter Birdwatch}). As for Youtube, this platform utilizes the \href{https://schema.org/ClaimReview}{schema.org ClaimReview} markup, where fact-checking articles created by eligible publishers can appear on information panels (see table \ref{summary} in Appendix \ref{links}).

@@

During the COVID19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see table \ref{summary} in Appendix \ref{links}). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. In particular, policies regarding misinformation are not part of the set of platform rules or community guidelines (as of July 2021). The 2019 report of the Facebook Data Transparency Advisory Group (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report.}''. Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on Youtube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

@@

%{\color{pink} Note, put here something along the following lines: cite misinformation review article about support for fact-checking + hard to correct beliefs once it goes viral. cite Pennycook / Rand? }
%In particular, determining the truthfulness of content over controversial topics or ongoing political events proves to be a tricky task and correcting beliefs about viral content, fact-checked as false ... 
%For an overview of Youtube's policy on fact-checking see: \href{https://support.google.com/youtube/answer/9229632}{support.google.com/youtube/answer/9229632}. To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} 
%In particular, the peculiar task of determining the truthfulness of content over controversial topics or ongoing political events 
%\smallskip

@@

\hline
&  \begin{tabular}[c]{@{}l@{}} Application Programming Interface (API)  \end{tabular}                                                                                                                                                                                                                                                                                                                                                                                                                                            &    \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}} Web Scraping \end{tabular}}                                                                                                                                                                 \\ \hline
 \includegraphics[scale=0.05]{./img/fb_logo.png} & \begin{tabular}[c]{@{}l@{}}\href{https://www.crowdtangle.com}{CrowdTangle API} and  \href{https://buzzsumo.com}{Buzzsumo API}   \end{tabular}                                                                                                                                                                                                               &      \multicolumn{1}{l|}{}    
\\ \hline
 \includegraphics[scale=0.007]{./img/tw_logo.png}  &  \href{https://developer.twitter.com/en/docs/twitter-api/early-access}{Twitter API V2} & \multicolumn{1}{l|}{\href{https://github.com/medialab/minet/blob/master/docs/cli.md\#twitter-scrape}{Minet tw scrape}}                                                                                                                                                                           \\ \hline
 \includegraphics[scale=0.03]{./img/yt_logo.png}  &  \href{https://developers.google.com/youtube/v3}{Youtube API V3}      &                                                                                                                            \multicolumn{1}{l|}{}                                                                                                                                                                           \\ \hline
\end{tabular}
\caption{Data collection}
\label{tab1}
\end{table}

Nevertheless, misinformation specific interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify with data mining mainstream platforms' current actions regarding content with low credibility or false information. We do so by providing a series of examples for different interventions and platforms. We only focus on three platforms: Facebook, Twitter and Youtube. Both Facebook and Youtube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. More specifically, we survey a number of common policies used in order to tackle misinformation, across the three above cited mainstream platforms, Facebook, Twitter and Youtube: temporary or permanent suspension of users, reducing the visibility of some content, introducing flags and notices. We do not provide  an exhaustive list of methods on how to investigate the platforms’ policies. We rather provide a methodology to investigate key policies, that can be useful to researchers or journalists interested in implementing external monitoring. For a detailed list of interventions of multiple platforms we invite the reader to check the following \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} compiled by Slatz and Leibowicz (2021)~\cite{niemanlab}.

@@

 We summarize in table \ref{tab1} the tools used to collect the data from Facebook, Twitter and Youtube, that we use in multiple examples throughout the present article. We compile in Appendix \ref{links} in table \ref{summary}, a list of links that redirect to the policies, regulations and transparency centers of Facebook, Twitter and Youtube that we discuss throughout the article. Finally, we discuss how an increased effort of transparency regarding specific content can help the community of researchers study and assess the impact of platforms' policies regarding misinformation.  %the announced policy for a specific user was implemented or not.

@@

\subsection{Youtube}

In this section, we turn to the channel's temporary or permanent suspension policy of Youtube. Whenever a channel publishes a video that violates the community guidelines for the first time they will  usually receive a warning and the content will be removed. For the second time the channel will start receiving strikes. A first strike results in limiting the access of the Youtube channel  for one week, like uploading videos, streaming and other activities. Then a second strike is similar but the suspension will be for two weeks. A third strike results in the termination of the channel. The strike count of a channel lasts 90 days.  In the special case, where a video is in extreme violation of the guidelines, the publishing channel may get terminated without a warning. To illustrate the implementation of this policy we provide two examples for the temporary suspension of the following two Youtube channels: \href{https://www.youtube.com/channel/UCNbIDJNNgaRrXOD7VllIMRQ}{One America news Network} and \href{https://www.youtube.com/user/TonyHeller1}{Tony Heller}.

@@

	\caption{Panel (a): Tweet announcing moving to rumble by OANN (Twitter), Twitter ID \href{https://twitter.com/OANN/status/1372238828425998336}{1372238828425998336}. Panel (b): Tony Heller's tweet after getting suspended from Youtube, Twitter ID \href{https://twitter.com/Tony\_Heller/status/1310703852769796097}{1310703852769796097}.}
	\label{fig2_oann}
\end{figure}

@@

First, we investigate the temporary suspension case of the Youtube Channel of {\it One America News channel}. This channel received a first strike on November 24, 2020 for the promotion of  a false cure for COVID-19.\footnote{See nbcnews,  \href{https://www.nbcnews.com/news/all/youtube-suspends-oann-violating-its-covid-19-policy-n1248845}{YouTube suspends OANN for violating its Covid-19 policy} nbcnews, Ahiza García-Hodges, 24 11 2020.} We collected the activity of the channel OANN (video counts, view counts) using  the Youtube API v3, between November 2020 and January 2021. For the video counts, we used the playlist endpoint to retrieve the videos uploaded with their publishing date and for the view count we used the IDs of the videos we had from the playlists and via the videos endpoint we retrieved the view counts on June 2021.%\footnote{See the Google documentation  \href{https://developers.google.com/youtube/v3/docs/videos/list}{https://developers.google.com/youtube/v3/docs/videos/list} and \href{https://developers.google.com/youtube/v3/docs/playlists/list}{https://developers.google.com/youtube/v3/docs/playlists/list}}

\begin{figure}[h]
	\centering
	%\begin{multicols}{1}
		\includegraphics[scale=0.27]{./img/oann/fig1_oann.png}
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.26]{./img/oann/fig2_oann.png} 
	%\end{multicols}
	\caption{panel (a): Number of Youtube videos uploaded each day by the youtube channel {\it One America news Network} November 1, 2020 and January 1, 2021. Panel (b): accumulated view counts for videos. The metrics correspond to the videos’  publishing date and the data is retrieved from the youtube API with the {\it playlists} and  {\it videos} endpoints. }
	\label{fig1_oann}
\end{figure}

In addition, as shown in figure \ref{fig1_oann} when comparing the month before the suspension from 2020/10/24 to 2020/11/24 and one month after from 2020/12/01 to 2021/01/01 it was found that the view count decreased by -73\% and the videos uploaded by -55\%. Besides that, OANN decided to move officially to Rumble on March 17, 2021 as announced on their Twitter account (see figure \ref{fig2_oann}) and their upload activity on their Youtube channel is close to zero since that announcement.

%\smallskip
\begin{figure}[h]
	\centering
		%\begin{multicols}{1}
			\includegraphics[scale=0.35]{./img/tony/fig1_tony.png}
		%\end{multicols}
		%\begin{multicols}{1}
			\includegraphics[scale=0.34]{./img/tony/fig2_tony.png}
		%\end{multicols}
	\caption{Panel (a): Number of Youtube videos uploaded each day by the Youtube channel {\it Tony Heller} between September 1, 2020 and November 15, 2020. Panel (b): accumulated view counts for videos uploaded by the same Youtube channel. The date corresponds to the videos’  publishing date. 
}
	\label{fig1_tony}
\end{figure}

We now turn to our second example, the temporary suspension of the Youtube channel Tony Heller. This channel got its first strike after posting a video about an anti-covid-lockdown doctor getting arrested (see screenshot in figure \ref{fig2_oann}). The suspension period was for one week from September 29 until October 5. We applied the same methods as in the previous example for the data collection. Figure \ref{fig2_oann} shows the daily number of videos uploaded by the channel. The suspension period can be observed clearly in the historical data of the channel.  observing the reach of the audience 
Figure 3: Tony heller tweet after getting suspended from youtube (Twitter)
using view counts one month before the suspension starting from 2020/08/28 to 2020/09/28 and one month after the suspension from 2020/10/05 to 2020/11/05 the channel witnessed a drop of view counts by $-69.5\%$ and the videos published in the channel were less by $-29\%$. This drop in views can show that the  suspension  period
may  have  a  good  impact  on  reducing  the  audience interest or reach to the channel.

@@

\subsection{Youtube}

Youtube may provide an information panel for videos with topics that are prone to misinformation like COVID-19, moon landing, and climate change. Information panels provide ressources concerning a potentially controversial topic from independent third party partners or authoritative ressources. Youtube states that these panels exist regardless of the point of view expressed in a given video. In addition, these panels are not yet available in all languages and countries. %(“Information panel giving topical context”)

To further study the assignment of information panels to videos, we compiled a list of $171$ Youtube videos, {\color{red}checked by independent fact-checkers} and marked as containing false or inaccurate information. We collected the information panels, when present, in each video by connecting to a US server and scraping the content of the web page. Four types of information panels were found in the list of visited videos as shown in figure \ref{fig9}:  information about COVID-19, Vaccines, Climate change and US elections.

@@

\subsection{Youtube}

@@

	\item Make point about the communication of platforms for their policy: 4R of Youtube, RRI of Facebook,

@@

	\item After a channel gets suspended or a video is deleted the data related to them is removed from the youtube API. This is a problem that affected the study to investigate the suspension reasons or the removal. For instance, the original list of videos from table 1 had more than 200 videos available in March 2021, however, by June 2021 30 videos got deleted from youtube because they contain content that is against the youtube guidelines and their data got removed from the API.

>>>>

Section $230$ in the United States Communications Decency Act provides immunity for website platforms against the content created by users. Similar regulations exist in the European Union via the E-commerce Directive (2000) in articles $12$ and $15$.\footnote{See section $4$ in Bayer (2019)~\cite{Bayer} for a comprehensive overview of the mentioned articles.} Nevertheless, there is growing pressure for Mainstream Social Media Platforms, such as Facebook, Twitter or YouTube, to moderate the available content. In particular, platforms take explicit actions when content is in violation of local laws in different jurisdictions, such as laws regarding defamation of a racial nature, dissemination of symbols from unconstitutional organizations, privacy protection, digital security, electoral laws. For example, Facebook reports having implemented a total of $64.7$ thousand content restrictions based on local law across all countries in 2020.\footnote{See Facebook Transparency Center, Content restrictions based on Local Law: \href{https://transparency.fb.com/data/content\-restrictions}{transparency.fb.com/data/content\-restrictions}. We summed the count of content restrictions over all countries reported in the table, for $H1$ and $H2$ of the year 2020.} %Google reports a total of $26$ thousand government requests to remove content from July 2020 to December 2020, among which $11.4$ thousand concerned YouTube.\footnote{See Google's Transparency report, government requests to remove content: \href{https://transparencyreport.google.com/government-removals/overview}{transparencyreport.google.com/government-removals/overview}.} Twitter reports having received $42.2$ thousand legal demands from third-parties from January to June 2020, and has responded by withholding $82$ thousand accounts and $3.1$ thousand tweets.\footnote{See Twitter Transparency website, Removal requests: \href{https://transparency.twitter.com/en/reports/removal-requests.html\#2020-jan-jun}{transparency.twitter.com/en/reports/removal-requests.html\#2020-jan-jun}.}

@@

Furthermore, mainstream platforms are increasingly engaging in editorial tasks by implementing targeted policies to insure that each platform's rules are not violated. Community guidelines of Facebook, Twitter and YouTube can be summarized in a handful of categories, regarding safety, privacy and authenticity; which include sub-categories such as violence, terrorism, child sexual exploitation, abuse, harassment, hateful conduct, suicide or self-harm, illegal or regulated goods and services, platform manipulation and spam (see table \ref{summary} in Appendix \ref{links}). While specific to each platform, the previously cited categories correspond in most cases to well defined concepts that fall into legal frameworks in many countries, unlike misinformation. The intricacies of constructing a legal framework for misinformation arises from the difficulty of identifying and qualifying a piece of online content as false or misleading, among an overwhelming quantity of daily produced content, without infringing existing laws.\footnote{See \href{https://www.poynter.org/ifcn/anti-misinformation-actions/}{A guide to anti-misinformation actions around the world} on the website of Poynter Institute.} In particular, a number of recent studies  point towards the idea that ``Fake News'' or disinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense. \footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.'' Furthermore, the European commission recognizes the spread of online disinformation as a problem and has put together in 2018 a \href{https://digital-strategy.ec.europa.eu/en/policies/code-practice-disinformation}{Code of practice on Disinformation}, which is a set of self-regulatory standards to fight disinformation.}

@@

Hence, in the present article, we focus on mainstream platforms' policies and interventions regarding content with low credibility or false information, commonly referred to as {\it Fake News} (see Lazer et al. (2018)~\cite{lazer}). The {\it Fake News} phenomenon is still ill-defined by the academic community, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience.  The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society, in the rare event that it goes viral. This gave rise to a set of heterogenous fact-checking policies across mainstream platforms. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network. Facebook uses a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''}} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation (see \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{Twitter Birdwatch}). As for YouTube, this platform utilizes the \href{https://schema.org/ClaimReview}{schema.org ClaimReview} markup, where fact-checking articles created by eligible publishers can appear on information panels (see table \ref{summary} in Appendix \ref{links}).

@@

During the COVID19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see table \ref{summary} in Appendix \ref{links}). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. In particular, policies regarding misinformation are not part of the set of platform rules or community guidelines (as of July 2021). The 2019 report of the Facebook Data Transparency Advisory Group (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report.}''. Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

@@

%{\color{pink} Note, put here something along the following lines: cite misinformation review article about support for fact-checking + hard to correct beliefs once it goes viral. cite Pennycook / Rand? }
%In particular, determining the truthfulness of content over controversial topics or ongoing political events proves to be a tricky task and correcting beliefs about viral content, fact-checked as false ... 
%For an overview of YouTube's policy on fact-checking see: \href{https://support.google.com/youtube/answer/9229632}{support.google.com/youtube/answer/9229632}. To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} 
%In particular, the peculiar task of determining the truthfulness of content over controversial topics or ongoing political events 
%\smallskip

@@

\hline
&  \begin{tabular}[c]{@{}l@{}} Application Programming Interface (API)  \end{tabular}                                                                                                                                                                                                                                                                                                                                                                                                                                            &    \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}} Web Scraping \end{tabular}}                                                                                                                                                                 \\ \hline
 \includegraphics[scale=0.05]{./img/fb_logo.png} & \begin{tabular}[c]{@{}l@{}}\href{https://www.crowdtangle.com}{CrowdTangle API} and  \href{https://buzzsumo.com}{Buzzsumo API}   \end{tabular}                                                                                                                                                                                                               &      \multicolumn{1}{l|}{}    
\\ \hline
 \includegraphics[scale=0.007]{./img/tw_logo.png}  &  \href{https://developer.twitter.com/en/docs/twitter-api/early-access}{Twitter API V2} & \multicolumn{1}{l|}{\href{https://github.com/medialab/minet/blob/master/docs/cli.md\#twitter-scrape}{Minet tw scrape}}                                                                                                                                                                           \\ \hline
 \includegraphics[scale=0.03]{./img/yt_logo.png}  &  \href{https://developers.google.com/youtube/v3}{YouTube API V3}      &                                                                                                                            \multicolumn{1}{l|}{}                                                                                                                                                                           \\ \hline
\end{tabular}
\caption{Data collection}
\label{tab1}
\end{table}

Nevertheless, misinformation specific interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify with data mining mainstream platforms' current actions regarding content with low credibility or false information. We do so by providing a series of examples for different interventions and platforms. We only focus on three platforms: Facebook, Twitter and YouTube. Both Facebook and YouTube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. More specifically, we survey a number of common policies used in order to tackle misinformation, across the three above cited mainstream platforms, Facebook, Twitter and YouTube: temporary or permanent suspension of users, reducing the visibility of some content, introducing flags and notices. We do not provide  an exhaustive list of methods on how to investigate the platforms’ policies. We rather provide a methodology to investigate key policies, that can be useful to researchers or journalists interested in implementing external monitoring. For a detailed list of interventions of multiple platforms we invite the reader to check the following \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} compiled by Slatz and Leibowicz (2021)~\cite{niemanlab}.

@@

 We summarize in table \ref{tab1} the tools used to collect the data from Facebook, Twitter and YouTube, that we use in multiple examples throughout the present article. We compile in Appendix \ref{links} in table \ref{summary}, a list of links that redirect to the policies, regulations and transparency centers of Facebook, Twitter and YouTube that we discuss throughout the article. Finally, we discuss how an increased effort of transparency regarding specific content can help the community of researchers study and assess the impact of platforms' policies regarding misinformation.  %the announced policy for a specific user was implemented or not.

@@

\subsection{YouTube}

In this section, we turn to the channel's temporary or permanent suspension policy of YouTube. Whenever a channel publishes a video that violates the community guidelines for the first time they will  usually receive a warning and the content will be removed. For the second time the channel will start receiving strikes. A first strike results in limiting the access of the YouTube channel  for one week, like uploading videos, streaming and other activities. Then a second strike is similar but the suspension will be for two weeks. A third strike results in the termination of the channel. The strike count of a channel lasts 90 days.  In the special case, where a video is in extreme violation of the guidelines, the publishing channel may get terminated without a warning. To illustrate the implementation of this policy we provide two examples for the temporary suspension of the following two YouTube channels: \href{https://www.youtube.com/channel/UCNbIDJNNgaRrXOD7VllIMRQ}{One America news Network} and \href{https://www.youtube.com/user/TonyHeller1}{Tony Heller}.

@@

	\caption{Panel (a): Tweet announcing moving to rumble by OANN (Twitter), Twitter ID \href{https://twitter.com/OANN/status/1372238828425998336}{1372238828425998336}. Panel (b): Tony Heller's tweet after getting suspended from YouTube, Twitter ID \href{https://twitter.com/Tony\_Heller/status/1310703852769796097}{1310703852769796097}.}
	\label{fig2_oann}
\end{figure}

@@

First, we investigate the temporary suspension case of the YouTube channel of {\it One America News channel}. This channel received a first strike on November 24, 2020 for the promotion of  a false cure for COVID-19.\footnote{See nbcnews,  \href{https://www.nbcnews.com/news/all/youtube-suspends-oann-violating-its-covid-19-policy-n1248845}{YouTube suspends OANN for violating its Covid-19 policy} nbcnews, Ahiza García-Hodges, 24 11 2020.} We collected the activity of the channel OANN (video counts, view counts) using  the YouTube API v3, between November 2020 and January 2021. For the video counts, we used the playlist endpoint to retrieve the videos uploaded with their publishing date and for the view count we used the IDs of the videos we had from the playlists and via the videos endpoint we retrieved the view counts on June 2021.%\footnote{See the Google documentation  \href{https://developers.google.com/youtube/v3/docs/videos/list}{https://developers.google.com/youtube/v3/docs/videos/list} and \href{https://developers.google.com/youtube/v3/docs/playlists/list}{https://developers.google.com/youtube/v3/docs/playlists/list}}

\begin{figure}[h]
	\centering
	%\begin{multicols}{1}
		\includegraphics[scale=0.27]{./img/oann/fig1_oann.png}
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.26]{./img/oann/fig2_oann.png} 
	%\end{multicols}
	\caption{panel (a): Number of YouTube videos uploaded each day by the youtube channel {\it One America news Network} November 1, 2020 and January 1, 2021. Panel (b): accumulated view counts for videos. The metrics correspond to the videos’  publishing date and the data is retrieved from the youtube API with the {\it playlists} and  {\it videos} endpoints. }
	\label{fig1_oann}
\end{figure}

In addition, as shown in figure \ref{fig1_oann} when comparing the month before the suspension from 2020/10/24 to 2020/11/24 and one month after from 2020/12/01 to 2021/01/01 it was found that the view count decreased by -73\% and the videos uploaded by -55\%. Besides that, OANN decided to move officially to Rumble on March 17, 2021 as announced on their Twitter account (see figure \ref{fig2_oann}) and their upload activity on their YouTube channel is close to zero since that announcement.

%\smallskip
\begin{figure}[h]
	\centering
		%\begin{multicols}{1}
			\includegraphics[scale=0.35]{./img/tony/fig1_tony.png}
		%\end{multicols}
		%\begin{multicols}{1}
			\includegraphics[scale=0.34]{./img/tony/fig2_tony.png}
		%\end{multicols}
	\caption{Panel (a): Number of YouTube videos uploaded each day by the YouTube channel {\it Tony Heller} between September 1, 2020 and November 15, 2020. Panel (b): accumulated view counts for videos uploaded by the same YouTube channel. The date corresponds to the videos’  publishing date. 
}
	\label{fig1_tony}
\end{figure}

We now turn to our second example, the temporary suspension of the YouTube channel Tony Heller. This channel got its first strike after posting a video about an anti-covid-lockdown doctor getting arrested (see screenshot in figure \ref{fig2_oann}). The suspension period was for one week from September 29 until October 5. We applied the same methods as in the previous example for the data collection. Figure \ref{fig2_oann} shows the daily number of videos uploaded by the channel. The suspension period can be observed clearly in the historical data of the channel.  observing the reach of the audience 
Figure 3: Tony heller tweet after getting suspended from YouTube (Twitter)
using view counts one month before the suspension starting from 2020/08/28 to 2020/09/28 and one month after the suspension from 2020/10/05 to 2020/11/05 the channel witnessed a drop of view counts by $-69.5\%$ and the videos published in the channel were less by $-29\%$. This drop in views can show that the  suspension  period
may  have  a  good  impact  on  reducing  the  audience interest or reach to the channel.

@@

\subsection{YouTube}

YouTube may provide an information panel for videos with topics that are prone to misinformation like COVID-19, moon landing, and climate change. Information panels provide ressources concerning a potentially controversial topic from independent third party partners or authoritative ressources. YouTube states that these panels exist regardless of the point of view expressed in a given video. In addition, these panels are not yet available in all languages and countries. %(“Information panel giving topical context”)

To further study the assignment of information panels to videos, we compiled a list of $171$ YouTube videos, fact-checked by Science Feedback between April 2019 and March 2021 and marked as containing false or inaccurate information. We collected the information panels, when present, in each video by connecting to a US server and scraping the content of the web page. Four types of information panels were found in the list of visited videos as shown in figure \ref{fig9}:  information about COVID-19, Vaccines, Climate change and US elections.

@@

\subsection{YouTube}

@@

	\item Make point about the communication of platforms for their policy: 4R of YouTube, RRI of Facebook,

@@

	\item After a channel gets suspended or a video is deleted the data related to them is removed from the YouTube API. This is a problem that affected the study to investigate the suspension reasons or the removal. For instance, the original list of videos from table 1 had more than 200 videos available in March 2021, however, by June 2021 30 videos got deleted from YouTube because they contain content that is against the YouTube guidelines and their data got removed from the API.
